```{r libraries, echo = FALSE, message = FALSE}
# load libraries and any relevant data here
library(tidyverse)
library(stringr)
library(lubridate)
library(ggpubr)
library(stats)

recs <- readRDS("../data/recs_2015_cleaned.RDS")
weather <- read.csv("../data/2015 USA Weather Data FINAL.csv")
region <- read.csv("../data/states.csv")
```
# Fundamental Exploratory Analysis

```{r cleaning weather, echo=FALSE, message=FALSE}
#Data Wrangling + Merging for weather data
remove_doublequotes <- function(input){
  str_remove_all(string = input, pattern ="\"")  
} #This function removes the redundant double quotes from a column in a dataframe

weather$State.abb <- remove_doublequotes(input = weather$State.abb)
weather$State.name <- remove_doublequotes(input = weather$State.name)
weather$Timestamp <- remove_doublequotes(input = weather$Timestamp)

weather <- weather %>% left_join(region, by = c("State.name" = "State", "State.abb" = "State.Code")) %>% #Merging weather and region data 
  select(State.name, Region, AvgTemp, MaxTemp, MinTemp) %>% 
  filter(!is.na(Region))

temp_region <- weather %>% group_by(Region) %>% #Create a data table on temperature in each region
  summarise(AvgTemp_region = mean(AvgTemp),
            AvgMaxTemp = mean(MaxTemp),
            AvgMinTemp = mean(MinTemp)) 
```

```{r temperature on total energy, echo=FALSE, message=FALSE}
#Investigating any correlation between temperature and total energy use (in BTU)
x <- temp_region %>% ggplot(aes(x = Region, y = AvgTemp_region))+
  geom_col(fill = "lightblue", width = 0.5)+
  theme_minimal()+
  labs(y = "Average Temperature (in F)")

y <- recs %>%
  ggplot(aes(x = region_name, y = TOTALBTU)) +
  geom_boxplot() +
  theme_minimal()+
  labs(x = "Region", y = "Total energy use (in BTU)")

main_plot <- ggarrange(x, y, ncol = 2, nrow = 1, align = "v")
main_plot
```

It is quite interesting how Southern states use less energy compared to Midwestern and Northeastern states despite its higher temperature. This might be due to higher use of natural gas, propane in colder states. It is also intriguing to look into housing efficiency differences between regions.

```{r temperature on total electricity and natural gases, echo=FALSE, message=FALSE}
#Investigating any correlation between temperature and total energy use (in BTU)
y1 <- recs %>%
  ggplot(aes(x = region_name, y = BTUEL)) +
  geom_boxplot() +
  theme_minimal()+
  labs(x = "Region", y = "Total electricity use (in BTU)")

main_plot1 <- ggarrange(x, y1, ncol = 2, nrow = 1, align = "v")
main_plot1

y2 <- recs %>%
  ggplot(aes(x = region_name, y = BTUNG)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Region", y = "Total natural gases use (in BTU)")

main_plot2 <- ggarrange(x, y2, ncol = 2, nrow = 1, align = "v")
main_plot2
```

Looking closer to the usage of electricity, the South unsurprisingly consumes the most. This is likely because of their uses of AC to keep their house cool during hot weather. Whereas, the other regions contributes to much less electricity since they distribute their share of energy usage to natural gases and propane.

In conclusion, the location influences the total usage of energy source due to regional different temperature characteristics. However, it would be interesting to look into how weather conditions affect the energy efficiency level of a house, i.e. harsh weather hazards might wear the houses' conditions out.

```{r move in intervals effect on energy effciency of houses, echo=FALSE, message=FALSE}
move_diff_effect <- recs %>% mutate(move_diff = OCCUPYYRANGE- YEARMADERANGE, years_diff = 10 * move_diff) #calculate the intervals between when the house is built and when inhabitants moved into it

move_diff_effect %>%
 ggplot(aes(x = as.character(years_diff), y = TOTALDOL)) +
 geom_boxplot() +
 labs(title = "Impact of the gap between the construction year of a house and the year a family moves in on energy expenses ", x = "Gap (years)", y = "Total dollars spent ($)")+
 theme_minimal()

move_diff_effect %>%
 ggplot(aes(x = as.character(years_diff), y = TOTALBTU)) +
 geom_boxplot() +
 labs(title = "Impact of the gap between the construction year of a house and the year a family moves in on energy usage ", x = "Gap (years)", y = "Total energy use (BTU)")+
 theme_minimal()
```

It seems like the bigger the gap between construction year and move-in year is, the less money a household have to pay for their energy. This might due to the fact that older houses are more likely to undergo upgrades, advancing their energy efficiency.

```{r move in intervals effect on energy effciency of houses in different regions, echo=FALSE, message=FALSE}
 move_diff_effect %>%
   ggplot(aes(x = as.character(years_diff), y = TOTALDOL)) +
   geom_boxplot() +
   labs(title = "Impact of the gap between the construction year of a house and the year a family moves in on energy expenses in different regions",  x = "Gap (years)", y = "Total dollars spent ($)")+
   facet_wrap(~ region_name) +
   theme_minimal()
```

All regions' older houses have to pay less for energy usages. Specifically, the western households are shown to pay the least money for their energy. More investigation on differences in house architecture and efficiency among U.S. regions would help us gain a closer look to this phenomenon.

```{r move in intervals effect on energy effciency of houses among different housing types, echo=FALSE, message=FALSE}
move_diff_effect %>%
  ggplot(aes(x = as.character(years_diff), y = TOTALDOL)) +
  geom_boxplot() +
  labs(title = "Impact of the gap between the construction year of a house and the year a family moves in on energy expenses among different housing types",  x = "Gap (years)", y = "Total dollars spent ($)")+
  facet_wrap(~ housing_type) +
  theme_minimal()
```

```{r Midwest on total usage of natural gas, echo=FALSE, message=FALSE}
move_diff_effect %>% filter(region_name == "Midwest") %>%
 ggplot(aes(x = yearmade, y = BTUNG)) +
 geom_boxplot() +
 labs(title = "Impact of the construction year on natural gas usage in the Midwest",  x = "Year Made", y = "Total natural gas usage (BTU)")+
 theme_minimal()

move_diff_effect %>% filter(region_name == "Midwest") %>%
 ggplot(aes(x = yearmade, y = BTUNG)) +
 geom_boxplot() +
 scale_x_discrete(guide  = guide_axis(angle = 90))+
 labs(title = "Impact of the construction year on natural gas usage in the Midwest",  x = "Year Made", y = "Total natural gas usage (BTU)")+
 facet_wrap(~ housing_type) +
 theme_minimal()

move_diff_effect %>% filter(region_name == "Midwest") %>%
 ggplot(aes(x = yearmade, y = BTUEL)) +
 geom_boxplot() +
 scale_x_discrete(guide  = guide_axis(angle = 90))+
 labs(title = "Impact of the construction year on natural gas usage in the Midwest",  x = "Year Made", y = "Total natural gas usage (BTU)")+
 facet_wrap(~ housing_type) +
 theme_minimal()
```

It is not surprising that modern houses in the Midwest have better heat encapsulation, resulting in the lower usage of natural gas for heating. We further investigate the which housing type would be the most efficient for natural gas. Apartment large building seems to work really efficiently. However, this might because nowadays, apartment usually install two-way AC, so taking into account total electricity use would be helpful.


```{r insulation and total electricity used, echo=FALSE, message=FALSE}
insulation_level <- c("Not insulated", "Poorly insulated", "Adequately insulated", "Well insulated")

recs %>% filter(ng_warm == "Yes" & elec_warm == "No") %>%  ggplot(aes(x = fct_relevel(factor(adq_insulation), insulation_level), y = BTUNG))+
 geom_boxplot()+
 scale_x_discrete(guide  = guide_axis(angle = 90))+
 facet_wrap(~ region_name)

recs %>% filter(elec_warm == "Yes", elec_cool == "Yes") %>% ggplot(aes(x = housing_type, y = BTUEL))+
 geom_boxplot()
```

This seems weird since households with no insulation seem to use less electricity to heat their house. Next move is to search in the data codebook to verify this.

```{r window/room/ relationship with total efficiency, echo=FALSE, message=FALSE}
recs %>% ggplot(aes(x = factor(TOTROOMS), y = TOTALBTU)) +
 geom_boxplot()

recs %>% mutate(btu_per_sqft = TOTALBTU/TOTSQFT_EN) %>% ggplot(aes(x = factor(TOTROOMS), y = btu_per_sqft)) +
 geom_boxplot()

recs %>% group_by(WINDOWS) %>% summarise(n = n())
```

The number of households that have more than 20+ windows are so small that it doensn't show much.

```{r high ceiling, echo=FALSE, message=FALSE}
recs %>% filter(!is.na(high_ceiling)) %>% mutate(btu_per_sqft = TOTALBTU/TOTSQFT_EN) %>% ggplot(aes(x = high_ceiling, y = TOTALBTU))+
 geom_boxplot()
```

It is understandable that household with high ceiling would use more total energy in BTU compared to ones with low ceiling and less area space. 

### Regional Differences
My hypothesis is region that has:
  1. better socioeconomic status, which is influenced by educational levels 
  2. newer houses with stricter building codes and standards (insulation for heating and cooling)
  3. moderate weather patterns, which leads to more energy efficiency
  4. growing emergence of energy-efficient technologies and eco-conscious
would use less energy. 

```{r Socioeconomic Status, echo=FALSE, message=FALSE}
#Create socioeconomic status data set 
socioeconomic_status <- recs %>% mutate(energy_persqft = TOTALBTU/TOTSQFT_EN) %>% select(region_name, adq_insulation, energy_persqft, hh_income, EDUCATION)

#Create household income and insulation level data set
income_insul <- socioeconomic_status %>% group_by(hh_income, adq_insulation) %>% summarise(n = n()) %>% mutate(per = n*100/sum(n))

income_level <- c("Less than $20,000", "$20,000 - $39,999", "$40,000 - $59,999", "$60,000 to $79,999", "$80,000 to $99,999", "$100,000 to $119,999", "$120,000 to $139,999", "$140,000 or more")

insulation_level <- c("Not insulated", "Poorly insulated", "Adequately insulated", "Well insulated")

#Insulation status among different household incomes
ggplot(income_insul, aes(x = fct_relevel(factor(hh_income), income_level), y = per, fill = fct_relevel(factor(adq_insulation), insulation_level))) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Insulation level among different household incomes", x = "Household Incomes", y = "Frequency", fill = "Insulation level")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

#Energy use per square foot among different household incomes
socioeconomic_status %>% ggplot(aes(x = fct_relevel(factor(hh_income), income_level), y= energy_persqft))+
  geom_boxplot()+
  labs(title = "Energy use per square foot among differnt household incomes", x = "Household Incomes", y = "Energy per square foot")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))



#Fit a linear regression model investigating the energy used per square foot and household income, insulation level, and educational level
summary(lm(formula = energy_persqft ~ hh_income*adq_insulation + EDUCATION,data = socioeconomic_status))
```

We observe the trend of higher income households having better insulation status. Thus, we want to fit a model examining the relationship between total energy used per square foot and educational level, household incomes, and its interaction with insulation status. We observe household with higher income and higher education will have better energy efficiency in their houses (p < 0.05). On the other hand, although the interaction between household incomes and their insulation level demonstrates a positive trend with total use of energy per square foot, p-values are generally bigger to 0.15, implying this correlation is not statistically significant.

```{r Linear regression for regions socioeconomic status, echo=FALSE, message=FALSE}
#Function that summarise linear regression models of each region's socioeconomic status
lm_region_socioeconomic_summary <- function(outcome, variables, region){
  mod_formular_str <- str_c(outcome, "~", str_c(variables, collapse = "+"))
  mod_form <- as.formula(mod_formular_str)
  
  lm <- lm(mod_form, socioeconomic_status %>% filter(region_name == region))
  summary(lm)
}

variables1 = c("hh_income", "EDUCATION")
outcome = "energy_persqft"

lm_region_socioeconomic_summary(outcome, variables1, "West")
lm_region_socioeconomic_summary(outcome, variables1, "Midwest")
lm_region_socioeconomic_summary(outcome, variables1, "South")
lm_region_socioeconomic_summary(outcome, variables1, "Northeast")
```

Breaking down into four regions, it is quite interesting that the relationship between higher income and higher energy efficiency per square foot is not statistically significant (p > 0.05). However, household that have income less than $20,000 tend to use more energy per square foot. Education level is still have a negative relationship with energy efficiency as the lower educational level is, the more energy that households use. This suggests more intervention at state levels to educate household with low income and low educational level on how to efficiently use their energy.

```{r Education level, echo=FALSE, message=FALSE}
#Energy use per square foot among different educational levels
socioeconomic_status %>% ggplot(aes(x = factor(EDUCATION), y = energy_persqft))+
  geom_boxplot()+
  labs(title = "Energy use per square foot among different educational levels", x = "Education Levels", y = "Energy per square foot")+
  theme_minimal()
```

```{r Number of household each category, echo=FALSE, message=FALSE}
#Create data set on households' made year in four regions
region_yearMade <- recs %>% group_by(region_name, YEARMADERANGE) %>% summarise(n = n()) %>% mutate(per = n*100/sum(n))

ggplot(region_yearMade, aes(x = region_name, y = per, fill = as.character(YEARMADERANGE))) +
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "Propotion of household's year origin by region", fill = "Year Made Range", x = "Region",y = "Percentage")
```

Northeast is the region with oldest houses. This makes sense due to the high density population with limited land to build modern houses. 

```{r Median household ages, echo=FALSE, message=FALSE}
#Overview of median year made of households grouped by each region and insulation level
recs %>% group_by(region_name, adq_insulation) %>% summarise(median_age = median(YEARMADERANGE), median_BTU = median(TOTALBTU/TOTSQFT_EN))
```

We observe similar trends of year range made of the houses and their insulation status. The more modern the house is, the better it gets insulated. However, only in the West, poorly insulated households seem to be built before not insulated households. This is quite interesting and can be due to some policy (more research later on this). I would think better insulated house would use less energy (in BTU). However, this is not necessarily true due the confounding effect of household sizes. Indeed, bigger houses regardless of its insulation status would be more likely to use more energy. Therefore, we calculated the median of energy used per square foot. The results show that our hypothesis is correct: better insulated households would use consume less energy per square foot. However, well insulated households in the Northeast tend to use a lot more energy per square foot compared to other regions. WHY? 

```{r, Regional energy use discrepances among different insulation levels, echo=FALSE, message=FALSE}
recs %>% ggplot(aes(x = region_name, y = TOTALBTU/TOTSQFT_EN)) +
  geom_boxplot()+
  facet_wrap(~ adq_insulation)+
  labs(title = "Regional energy use discrepances among different insulation levels", x = "Region", y = "Energy per square foot")+
  theme_minimal()
```

Among four insulation statuses, Northeast uses the most energy per square foot. This can be due to higher residential density, resulting in more multifamily buildings than detached houses. Also, energy-related behavioral factors might play an important role in this trend.

```{r energy-related behavioral, echo=FALSE, message=FALSE}
remove_na <- function(data) {
  return(data[complete.cases(data), ]) # Function that remove rows with NA values from every column
}

energy_behavior <- recs %>% mutate(energy_persqft = TOTALBTU/TOTSQFT_EN) %>% select(region_name, TEMPGONE, TEMPHOME, TEMPNITE, summer_TEMPGONE, summer_TEMPHOME, summer_TEMPNITE, equip_used_short, energy_persqft) %>% remove_na()


#Function that summarise linear regression models of each region's households energy-related behavior
lm_region_summary <- function(outcome, variables, region){
  mod_formular_str <- str_c(outcome, "~", str_c(variables, collapse = "+"))
  mod_form <- as.formula(mod_formular_str)
  
  lm <- lm(mod_form, energy_behavior %>% filter(region_name == region))
  summary(lm)
}

variables = c("TEMPGONE", "TEMPHOME", "TEMPNITE", "summer_TEMPGONE", "summer_TEMPHOME", "summer_TEMPNITE", "equip_used_short")

lm_region_summary(outcome, variables, "West")
lm_region_summary(outcome, variables, "South")
lm_region_summary(outcome, variables, "Midwest")
lm_region_summary(outcome, variables, "Northeast")
```

Fitting models of energy-related behaviors among regions, we don't see any statistical significant trend between total used of energy per square foot and temperature set during day, night in winter and summer in the Midwest and Northeast. On the other hand, in the South, during winter, if houses increase their temperature by 1 degree F during winter, they will increase 0.43 BTU energy per square foot. In the summer, if houses increase their temperature by 1 degree F, they will decrease 0.44 BTU energy per square foot. In the West, during winter, keeping temperature low when leaving the house would save around 0.35 BTU energy per square feet, whereas increasing temperature during home and at night results in higher energy usage. Furthermore, in the South, households that can manually adjust or program thermostat are more energy efficient. In the West, unabling to adjust households temperature causes more energy loss.

```{r weather-related behavior, echo=FALSE, message=FALSE}
#Create data set on regional temperature
region_temp <- recs %>% group_by(region_name) %>% summarise(energy_persqft = sum(TOTALBTU)/sum(TOTSQFT_EN)) %>% left_join(temp_region, by = c("region_name" = "Region"))

region_temp %>% 
  ggplot(aes(x = region_name)) +
  geom_point(aes(y = energy_persqft), shape = 18, color = "forestgreen",
             size = 3) + #First y-axis on the total energy use per square foot
  geom_point(aes(y = AvgTemp_region), color = "lightblue") + #Second y-axis on the average temperature in F
  scale_y_continuous(
    name = "Energy use per square foot (BTU)",
    sec.axis = sec_axis(~.*1, name="Average temperature (Farenheit)")
  ) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45),
        axis.title.y = element_text(color = "forestgreen", size = 9),
        axis.title.y.right = element_text(color = "lightblue", angle = 90),
        axis.text.y = element_text(color = "forestgreen"),
        axis.text.y.right = element_text(color = "lightblue")) +
  labs(title = "Energy Use Disparities Across Region", x = "Region") 
```

Compared between different regions, North East uses most energy per square foot despite a milder weather compared to the Midwest and the South. Despite having the highest temperature, the South uses the second-least energy in BTU. This might due to better housing architect, allowing fresh air to flow through the household.